
```{r}
library(dplyr)
```
```{r}
library(tidyverse)
```

# Exploratory Data Analysis

### Importing dataset
```{r}
df <- read.csv("data_diabetes.csv", header = TRUE, sep = ",")
```

### Looking dataset
```{r}
dim(df)
```
```{r}
head(df)
```
```{r}
str(df)
```

```{r}
BasicSummary <- function(df, dgts=3){
  m <- ncol(df)
  varNames <- colnames(df)
  varType <- vector("character",m)
  topLevel <- vector("character",m)
  topCount <- vector("numeric",m)
  missCount <- vector("numeric",m)
  levels <- vector("numeric",m)
  
  for(i in 1:m){
    x<-df[,i]
    varType[i] <- class(x)
    xtab <- table(x,useNA="ifany")
    levels[i] <- length(xtab)
    nums <- as.numeric(xtab)
    maxnum <- max(nums)
    topCount[i] <- maxnum
    maxIndex <- which.max(nums)
    lvls <- names(xtab)
    topLevel[i] <- lvls[maxIndex]
    missIndex <- which((is.na(x)) | (x== "") | (x == " "))
    missCount[i] <- length(missIndex)
  }
  
  n <- nrow(df)
  topFrac <- round(topCount/n, digits = dgts)
  missFrac <- round(missCount/n, digits = dgts)
  summaryFrame <- data.frame(variable = varNames, type = varType, levels = levels, topLevel = topLevel, topCount = topCount, topFrac = topFrac, missFreq = missCount, missFrac = missFrac)
  return(summaryFrame)
}
```

```{r}
BasicSummary(df)
```
```{r}
summary(df)
```
1. 100000 observations with 9 attributes.
2. Three is three data type: integer, character, numeric.
3. Integer data type: hypertension, heart_disease, blood_glucose_level, diabetes.
4. Character data type: gender and smoking_history.
5. Numeric data type: age, bmi, and HbA1c_level.
6. There are no missing value in every variable


## Checking Duplicates
```{r}
duplicate_count <- sum(duplicated(df))
print(duplicate_count)
```
```{r}
duplicate_rows <- duplicated(df)

df <- df[!duplicate_rows, ]
```
```{r}
dim(df)
```
There are 3854 duplicates removed.
The reason we removed the duplicate is to reduce the amount of inaccuracies

We'll be looking at outliers in age, bmi, HbA1c_level, and blood_glucose_level
```{r}
meanAge <- mean(df$age)
sdAge <- sd(df$age)
par(mfrow=c(1,2), pty="m")

plot(df$age, main='Three Sigma Edit Rule', xlab='Record',ylab='age',ylim=c(-50,150))
abline(h=meanAge+3*(sdAge),col='purple',lwd=3,lty='dotdash')
abline(h=meanAge-3*(sdAge),col='purple',lwd=3,lty='dotdash')

boxplot(df$age, main = "Boxplot of age", ylab = "Value")
outliers <- boxplot.stats(df$age)$out
points(rep(1, length(outliers)), outliers, col = "red", pch = 16)
legend("topright", legend = "Outliers", col = "red", pch = 16)
```
```{r}
meanHb <- mean(df$HbA1c_level)
sdHb <- sd(df$HbA1c_level)
par(mfrow=c(1,2), pty="m")

plot(df$HbA1c_level, main='Three Sigma Edit Rule', xlab='Record',ylab='HbA1c',ylim=c(2,10))
abline(h=meanHb+3*(sdHb),col='purple',lwd=3,lty='dotdash')
abline(h=meanHb-3*(sdHb),col='purple',lwd=3,lty='dotdash')

boxplot(df$HbA1c_level, main = "Boxplot of HbA1c_level", ylab = "Value")
outliers <- boxplot.stats(df$HbA1c_level)$out
points(rep(1, length(outliers)), outliers, col = "red", pch = 16)
legend("topright", legend = "Outliers", col = "red", pch = 16)
```

```{r}
meanBlood <- mean(df$blood_glucose_level)
sdBlood <- sd(df$blood_glucose_level)
par(mfrow=c(1,2), pty="m")

plot(df$blood_glucose_level, main='Three Sigma Edit Rule', xlab='Record',ylab='Blood glucose level',ylim=c(0,350))
abline(h=meanBlood+3*(sdBlood),col='purple',lwd=3,lty='dotdash')
abline(h=meanBlood-3*(sdBlood),col='purple',lwd=3,lty='dotdash')

boxplot(df$blood_glucose_level, main = "Boxplot of blood glucose level", ylab = "Value")
outliers <- boxplot.stats(df$blood_glucose_level)$out
points(rep(1, length(outliers)), outliers, col = "red", pch = 16)
legend("topright", legend = "Outliers", col = "red", pch = 16)
```

1. There are no outliers in the age variable.
2. There are some outliers in the HbA1c_level and blood glucose level, but this is a health related dataset, there are bound to be rare value in health related dataset.

```{r}
par(mfrow = c(1,3))
qqnorm(df$age,main="Age Normality")
qqline(df$age)

qqnorm(df$HbA1c_level,main="HbA1c Normality")
qqline(df$HbA1c_level)

qqnorm(df$blood_glucose_level,main="Blood glucose level Normality")
qqline(df$blood_glucose_level)
```
##EDA

```{r}
plot(df$age, df$heart_disease, main="Relation between age and heart disease")
```

```{r}
plot(df$bmi, df$heart_disease, main="Relation between BMI and Heart disease")
```

```{r}
plot(df$age, df$diabetes, main="Relation between age and diabetes")
```

```{r}
plot(df$bmi, df$diabetes, main="Relation between BMI and diabetes")
```

```{r}
plot(df$blood_glucose_level, df$diabetes, main="Relation between blood glucose level and diabetes")
```

```{r}
plot(df$HbA1c_level, df$diabetes, main="Relation between HbA1c level and diabetes")

```
Based on the data obtained above, age has an effect on heart problems, but has no effect on diabetes. The BMI of people who have heart problems and diabetes have a nearly identical range, between 10 and 60. People who have diabetes have higher blood glucose levels and HbA1c levels than those who don't have diabetes.


##Data anomalies
```{r}
freq_table2 <- table(df$smoking_history)
freq_table2
```
Because there are a lot of "no info" part on the smoking_history, it is better to drop the column.

```{r}
#Dropping smoking_history table
df <- df[, -5]
head(df)
```
```{r}
freq_table1 <- table(df$gender)
freq_table1
```
We want to encode the gender into 0 and 1
Because the "Other" gender is not specified it is better to drop evey record with the gender "other"

```{r}
df <- df[df$gender != 'Other', ]
```
```{r}
freq_table1 <- table(df$gender)
freq_table1
```
encode the gender column into 0 and 1

```{r}
df <- df %>% 
  mutate(gender = recode(gender, "Female" = 0, "Male" = 1))
head(df)
```
```{r}
freq_table1 <- table(df$gender)
freq_table1
```

```{r}
str(df)
```
all the variable is in the right data type

## Correlation between variables
```{r}
library(corrplot)
cor_matrix <- cor(df)
corrplot(cor_matrix, method = "color")
```
```{r}
print(cor_matrix)
```
From this pearson correlation, as we can see the highest correlation between diabetes and blood_glucose_level (0.42436644). With the second highest is between diabetes and HbA1c_level (0.40644594)

# Model
```{r}
library(mlbench)
library(dplyr)
library(rpart)
library(rpart.plot)
library(pROC)
library(MLmetrics)
```

## Logistic Regression
## Splitting dataset
```{r}
set.seed(123)

n <- nrow(df)
train <- sample(n, round(0.7 * n))
DiaTrain <- df[train,]
DiaValidation <- df[-train,]
head(DiaTrain)
head(DiaValidation)
dim(DiaTrain)
dim(DiaValidation)
```
## Model Fitting

For the first model we are using all the variable
```{r}
logisticFull <- glm(diabetes ~., family="binomial", data = df)
summary(logisticFull)
```
all the variables are highly statistically significant.

According to cdc people are at risk of type 2 diabetes if: have prediabetes, are overweight, are 45 years or older, have family relatives with type 2 diabetes, and physically not active.
Thus, for the second model We choose some selected variable such as bmi, age, and HbA1c_level.
```{r}
logistic1 <- glm(diabetes ~ age + bmi + HbA1c_level, family="binomial", data = df)
summary(logistic1)
```

```{r}
phatFull <- predict(logisticFull, newdata = df, type = "response")
phatSelect <- predict(logistic1, newdata = df, type = "response")
```

## Performance (logsitic)
### ROC
```{r}
ROCFull <- roc(df$diabetes, phatFull, plot=TRUE, print.auc = TRUE)
```
```{r}
ROCSelect <- roc(df$diabetes, phatSelect, plot=TRUE, print.auc = TRUE)
```
### AUC
```{r}
AUC(phatFull, DiaValidation$diabetes)
```
```{r}
AUC(phatSelect, DiaValidation$diabetes)
```

### Confusion Matrix & Accuracy
```{r}
threshold <- table(df$diabetes, phatFull > 0.5)
accuracy <- round(sum(diag(threshold)) / sum(threshold), 2)
threshold
sprintf("Accuracy is %s", accuracy)
```

```{r}
threshold <- table(df$diabetes, phatSelect > 0.5)
accuracy <- round(sum(diag(threshold)) / sum(threshold), 2)
threshold
sprintf("Accuracy is %s", accuracy)
```
# Conclusion
For this logistic regression, We are using two kind of model, the first model with all the variable and the second model with age, bmi, and HbA1c_level as the variable.
The first model that we were using are statistically high in significant and the secind model are too.
The result show there are higher accuracy in the first model with 0.96 accuracy rather than the second model with 0.94 accuracy. Thus, the first model with all variable are outperforming second variable with three selected variable.



